Substructure Boosts to Dark Matter Annihilation from Sommerfeld Enhancement

### 2009-02-26 tex/somm.tex
Improvemenets after Neal's comments

Of course, when considering the Sommerfeld enhancement of a halo of a
certain velocity dispersion $\sigv$ we need to average the Sommerfeld
enhancement over the distribution of relative velocities of that
halo. We will approximate the one-particle velocity distribution of
the halo as a single truncated Maxwell-Boltzmann distribution with
velocity dispersion $\sigv$
\begin{equation}\label{eq:maxboltz}
f(v) \propto \left\{
\begin{array}{rl}
v^2\, e^{-v^2/2\sigv^2}  & \qquad v \leq v_{\mathrm{esc}}\\
0 & \qquad v > v_{\mathrm{esc}}\, ,\\
\end{array} \right.
\end{equation}
in which we set the escape velocity to $v_{\mathrm{esc}} =
\sqrt{2}\,\sigv$. This approximation of a single Maxwell-Boltzmann
distribution for the velocity distribution means that we assume that
the halo has a constant velocity dispersion over the halo, \ie, that
the halo is isothermal. This is not necessarily the case ---indeed, it
is not if we believe that the density profiles of \DM\ halos are well
fit by Einasto or \NFW\ profiles--- but it is not a bad assumption for
intermediate distances from the halo's center, \ie, between the
innermost and outermost parts of the halo
\cite{2007ApJ...667L..53W}. A more detailed treatment of the
Sommerfeld enhancement in subhalos should take the non-uniform
velocity dispersion in subhalos into account. Note that in the pure
$1/v$ regime the averaging will tend to sustain the $1/v$ behavior up
to a factor $\order(1)$, as we have that
\begin{equation}
\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2}
\somm(v)\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2} \right.\propto
\frac{1}{\sigv}\, ,
\end{equation}
since the truncation does not significantly influence the
result. Similarly, in the resonance region the $1/v^2$ behavior is
conserved by the averaging. When the enhancement has leveled off the
averaging has no effect.

@@

\begin{figure*}[t]
\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 488
\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 488
\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 488
\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 488
\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 488
\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 488
\caption{Predicted \DM\ annihilation fluxes for various dwarf
Spheroidal satellite galaxies. The properties of the \dSphs\ used are
given in Table \ref{tab:dsphs}.}%
\label{fig:dsph}%
\end{figure*}

@@

{\bf Acknowledgements:} It is a pleasure to thank Mulin Ding and David
Hogg for comments and assistance. This research was partially
supported by NASA (grant NNX08AJ48G).

>>>>

Of course, when considering the Sommerfeld enhancement of a halo of a
certain velocity dispersion $\sigv$ we need to average the Sommerfeld
enhancement over the distribution of relative velocities of that
halo. We will approximate the one-particle velocity distribution of
the halo as a single truncated Maxwell-Boltzmann distribution with
velocity dispersion $\sigv$
\begin{equation}\label{eq:maxboltz}
f(v) \propto \left\{
\begin{array}{rl}
v^2\, e^{-v^2/2\sigv^2}  & \qquad v \leq v_{\mathrm{esc}}\\
0 & \qquad v > v_{\mathrm{esc}}\, .\\
\end{array} \right.
\end{equation}
This approximation of a single Maxwell-Boltzmann
distribution for the velocity distribution means that we assume that
the halo has a constant velocity dispersion over the halo, \ie, that
the halo is isothermal. This is not necessarily the case ---indeed, it
is not if we believe that the density profiles of \DM\ halos are well
fit by Einasto or \NFW\ profiles--- but it is not a bad assumption for
intermediate distances from the halo's center, \ie, between the
innermost and outermost parts of the halo
\cite{2007ApJ...667L..53W}. A more detailed treatment of the
Sommerfeld enhancement in subhalos should take the non-uniform
velocity dispersion in subhalos into account. Note that in the pure
$1/v$ regime the averaging will tend to sustain the $1/v$ behavior up
to a factor $\order(1)$, as we have that
\begin{equation}
\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2}
\somm(v)\left/\int_0^\infty \dd v \, v^2\, e^{-v^2/4\sigv^2} \right.\propto
\frac{1}{\sigv}\, ,
\end{equation}
since the truncation does not significantly influence the
result. Similarly, in the resonance region the $1/v^2$ behavior is
conserved by the averaging. When the enhancement has leveled off the
averaging has no effect.

@@

\begin{figure*}[t]
\includegraphics[clip=]{flux_carina_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_draco_65.eps}\\%%BoundingBox: 190 290 433 500
\includegraphics[clip=]{flux_fornax_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_sextans_65.eps}\\%%BoundingBox: 190 290 433 500
\includegraphics[clip=]{flux_umaii_65.eps}%%BoundingBox:  171 290 412 500
\includegraphics[clip=]{flux_ursaminor_65.eps}%%BoundingBox: 190 290 433 500
\caption{Predicted \DM\ annihilation fluxes for various dwarf
Spheroidal satellite galaxies. The properties of the \dSphs\ used are
given in Table \ref{tab:dsphs}.}%
\label{fig:dsph}%
\end{figure*}

@@

{\bf Acknowledgements:} It is a pleasure to thank Mulin Ding, David
Hogg, and Neal Weiner for comments and assistance. This research was partially
supported by NASA (grant NNX08AJ48G).
